%----------------------------------------------------------------------------------------
%	COPYRIGHT PAGE
%----------------------------------------------------------------------------------------

\makeatletter
\uppertitleback{\@titlehead} % Header

\lowertitleback{
	% \textbf{Disclaimer}\\
	% You can edit this page to suit your needs. For instance, here we have a no
	% copyright statement, a colophon and some other information. This page is
	% based on the corresponding page of Ken Arroyo Ohori's thesis, with minimal
	% changes.
	
	% \medskip
	
	\textbf{Some rights reserved}\\
	\ccby\ This work is available under the Creative Commons Attribution 3.0 IGO
	license (CC BY 3.0 IGO) \href{http://
	creativecommons.org/licenses/by/3.0/igo}{http://
	creativecommons.org/licenses/by/3.0/igo}. Under the Creative Commons
	Attribution license, you are free to copy, distribute, transmit, and adapt
	this work, including for commercial purposes, under the following
	conditions:

	\begin{itemize}
		\item Attribution—Please cite the work as follows: Serrano Juan Miguel,
		Roca Lidia , Palenzuela Patricia. \@date. \@title~
		DOI:~\doiHref. License: Creative Commons Attribution CC
		BY 3.0 IGO  
		\item Translations—If you create a translation of this work, please add
		the following disclaimer along with the attribution: This translation
		was not created by \shortauthor~ and should not be considered an official
		\shortauthor~ translation. \shortauthor~ shall not be liable for any content
		or error in this translation.  
		
		\item Adaptations—If you create an adaptation of this work, please add
		the following disclaimer along with the attribution: This is an
		adaptation of an original work by \shortauthor. Views and opinions
		expressed in the adaptation are the sole responsibility of the author or
		authors of the adaptation and are not endorsed by \shortauthor.  
		
		\item Third-party content—\shortauthor~ does not necessarily own each
		component of the content contained within the work. \shortauthor~
		therefore does not warrant that the use of any third-party-owned
		individual component or part contained in the work will not infringe on
		the rights of those third parties. The risk of claims resulting from
		such infringement rests solely with you. If you wish to re-use a
		component of the work, it is your responsibility to determine whether
		permission is needed for that re-use and to obtain permission from the
		copyright owner. Examples of components can include, but are not limited
		to, tables, figures, or images.
	\end{itemize}
		
	\medskip

	\textbf{Available at} \\
	DOI: \doiHref \\
	Main repository: \url{\repositoryBaseUrl} \\
	Mirror repository: \url{\repositoryMirrorUrl}\\
	
	\medskip
	
	\textbf{Colophon} \\
	This document was typeset with the help of
	\href{https://sourceforge.net/projects/koma-script/}{\KOMAScript} and
	\href{https://www.latex-project.org/}{\LaTeX} using the
	\href{https://github.com/fmarotta/kaobook/}{kaobook} class.

	The source code of this book class is available at:~\url{https://github.com/fmarotta/kaobook}

	\medskip
	
	\textbf{Publisher} \\
	First printed in \@date~by \@publishers
}
\makeatother